---
title: 'Statistical report: Phenotyping heart failure patients with iron deficiency/amenia for iron treatment and health care outcomes'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB3, https://kiheartfailure.github.io/shfdb3/, v 3.1.5. 

## Inclusion/exclusion criteria

Two populations are used, one for the main analysis, A/ID status and FCM use, 
the latter on a subset, 
and one for the distribution of recorded measures for transferrin/ferritin (TF population).

```{r flow}
mykable(flow, caption = "Flowchart")
```

First patient in: `r min(pdata$shf_indexdtm)` and last patient in: `r max(pdata$shf_indexdtm)`.  

The median age (IQR) is `r pdata %>% summarise(med = dF(median(shf_age), dig = 1),
                                             q1 = dF(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = dF(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r pdata %>% count(shf_sex) %>%
  mutate(perc = dF(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    

```{r flowtf}
mykable(flow_tf, caption = "Flowchart for TF population")
```

First patient in for the TF population: `r min(pdata_tf$shf_indexdtm)` and last patient in: `r max(pdata_tf$shf_indexdtm)`.  

The median age (IQR) in the TF population is `r pdata_tf %>% summarise(med = dF(median(shf_age), dig = 1),
                                             q1 = dF(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = dF(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r pdata_tf %>% count(shf_sex) %>%
  mutate(perc = dF(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    
  
## Created variables 

```{r npr}
mykable(metaout, caption = "Additional outcomes from NPR")
```

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

# Transferrin/ferritin measure

## Missing data

Missing data was imputed with multiple imputation (n = 10) using mice [@mice]. 
Variables included in the model are indicated in 
Table \ref{tab:tab1tf}. The primary outcome, 
first heart failure hospitalization/all-cause death,
was included as the Nelson-Aalen estimator. Anemia and FCM are included 
in the imputation model but not transferrin/ferritin. 

## Baseline characteristics according to measure of transferrin/ferritin

The baseline characteristics are presented partly for 4 groups 
(measure on neither transferrin nor ferritin, measure on only transferrin, only on ferritin and on both)
and partly for two groups (measure on either transferrin or ferritin or measure on none). 
All other analysis on transferrin/ferritin measure are performed on the latter. 

```{r, child = "../src/tab1_tf.Rmd"}

```

```{r, child = "../src/geo_tf.Rmd"}

```

\clearpage
\newpage

## Predictors of transferrin/ferritin measure

The association between baseline covariates and transferrin/ferritin measure
was investigated using logistic regression. 

```{r, child = "../src/pred_tf.Rmd"}

```

### Assumptions

Outliers were investigated with Cook's distance and multicollinearity 
with the variance inflation factor in the full model in the overall population.
=> No action deemed necessary. 

\clearpage
\newpage

# A/ID status

## Missing data

Missing data was imputed with multiple imputation (n = 10) using mice [@mice]. 
Variables included in the model are indicated in 
Table \ref{tab:tab1aidall}. The primary outcome, 
first heart failure hospitalization/all-cause death,
was included as the Nelson-Aalen estimator. Anemia, ID and FCM are not included 
in the imputation model. 

## Baseline characteristics according to A/ID status

```{r, child = "../src/tab1_aid.Rmd"}

```

## Outcome analysis according to the A/ID status

The following outcomes are considered: 

- All-cause mortality/first HF hospitalization (primary endpoint)
- All-cause mortality
- First HF hospitalization
- First all-cause hospitalization
- First HF visit

- Repeated HF hospitalization
- Repeated all-cause hospitalization
- Repeated HF visit

Data were censored at 2018-12-31 or death/emigration. 

The outcomes are presented with the 1 - Kaplan-Meier curves with the overall p 
from the crude cox regression presented in the figure. Cox proportional hazards regressions were 
used to model the time to first event analysis and negative 
binomial regression including the log of time as an offset in the model for 
the repeated events. The models were partly crude and partly adjusted for variables 
indicated in Table \ref{tab:tab1aidall}.

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the primary outcome. 
Possible outliers were visually inspected by plotting the dfbetas. 
=> No action deemed necessary. 

The median (min-max) follow-up is 
`r pdata %>% summarise(med = dF(median(sos_outtime_death / 365.25 * 12), dig = 1),
                                             min = dF(min(sos_outtime_death / 365.25 * 12), dig = 1),
                                             max = dF(max(sos_outtime_death / 365.25 * 12), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r pdata %>% summarise(sumpy = dF(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.


```{r, child = "../src/km_aid.Rmd"}

```

```{r, child = "../src/outtab_aid.Rmd"}

```

\clearpage
\newpage

# FCM 

In this analysis patients are included if the are ID 
(ferritin <100 µg/L, or ferritin between 100–299 µg/L and transferrin saturation <20%) or
are taking FCM.

## Missing data

The same imputation was used as for A/ID.

## Propensity scores and matching

A propensity score for treatment with FCM was estimated for each patient with 
logistic regression for each of the 10 imputed datasets using the variables 
indicated in Table \ref{tab:tab1fcm}. 1:2 matching without replacement [@match] 
was thereafter performed on the average of the resulting 10 ps [@psmatch]. Matching was 
allowed if the propensity score differed by 0.01 or less. The ability of the ps 
matching to balance the baseline characteristics was assessed by 
standardized mean differences.  

1:2 has deemed the best option when taking into account the number of patients retained and balance between groups.
Other matching ratios: `r matchingn`.

## Baseline characteristics according to the use of FCM 

```{r, child = "../src/tab1_fcm.Rmd"}

```

```{r, child = "../src/densitydose_fcm.Rmd"}

```

```{r, child = "../src/geo_fcm.Rmd"}

```

\clearpage
\newpage

## Outcomes according to the use of FCM

Data were censored at 2018-12-31 or death/emigration. 

The outcomes are presented with the 1 - Kaplan-Meier curves with the p 
from the cox regressions presented in the figure. Cox proportional hazards regressions were 
used to model the time to first event analysis and negative 
binomial regression including the log of time as an offset in the model for 
the repeated events. The models were partly crude, partly adjusted for 
the individual variables indicated in Table \ref{tab:tab1fcm} and partly in the matched cohort, 
modelling the dependence between pairs with a strata term. 

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the primary outcome 
in the ps matched cohort. 
Possible outliers were visually inspected by plotting the dfbetas. 
=> No action deemed necessary. 

The median (min-max) follow-up is 
`r pdata_fcm %>% summarise(med = dF(median(sos_outtime_death / 365.25 * 12), dig = 1),
                                             min = dF(min(sos_outtime_death / 365.25 * 12), dig = 1),
                                             max = dF(max(sos_outtime_death / 365.25 * 12), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r pdata_fcm %>% summarise(sumpy = dF(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.


```{r, child = "../src/km_fcm.Rmd"}

```

```{r, child = "../src/outtab_fcm.Rmd"}

```

\clearpage
\newpage

### Subgroups

Performed using an interaction term between FCM and EF on the patients in the matched cohort. 
Note that there is missing data for EF so the matching (and thereby the adjustment) 
may therefore be incomplete. 

```{r, child = "../src/outtab_fcm_ef.Rmd"}

```

\clearpage
\newpage

## Predictors of FCM use 

The association between baseline covariates and FCM was investigated using logistic regression. 

```{r, child = "../src/pred_fcm.Rmd"}

```

### Assumptions

Outliers were investigated with Cook's distance and multicollinearity 
with the variance inflation factor in the full model in the overall population.
=> No action deemed necessary. 

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/iron. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
