---
title: 'Statistical report: Phenotyping heart failure patients with iron deficiency/amenia for iron treatment and health care outcomes'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoffigures
\listoftables
\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB3, https://kiheartfailure.github.io/shfdb3/, v 3.1.5. 

## Inclusion/exclusion criteria

```{r flow}
mykable(flow, caption = "Flowchart")
```

First patient in: `r min(pdata$shf_indexdtm)` and last patient in: `r max(pdata$shf_indexdtm)`.  

The median age (IQR) is `r pdata %>% summarise(med = dF(median(shf_age), dig = 1),
                                             q1 = dF(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = dF(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r pdata %>% count(shf_sex) %>%
  mutate(perc = dF(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    
  
## Created variables 

```{r npr}
mykable(metaout, caption = "Additional outcomes from NPR")
```

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Missing data was imputed with multiple imputation (n = 10) using mice [@mice]. 
Variables included in the model are indicated in 
Table \ref{tab:tab1aid}. The primary outcome, 
first heart failure hospitalization/all-cause death,
was included as the Nelson-Aalen estimator. 

## Baseline characteristics according to A/ID status

```{r, child = "../src/tab1_aid.Rmd"}

```

## Outcome analysis according to the A/ID status

The following outcomes are considered: 

- All-cause mortality/first HF hospitalization (primary endpoint)
- All-cause mortality
- First HF hospitalization
- First all-cause hospitalization
- First HF visit

- Repeated HF hospitalization
- Repeated all-cause hospitalization
- Repeated HF visit

Data were censored at 2018-12-31 or death/emigration. 

The outcomes are presented with the 1 - Kaplan-Meier curves
and tested with the log-rank test. Cox proportional hazards regressions were 
used to model the time to first event analysis and negative binomial regression to model 
the repeated events. The models were partly crude and partly adjusted for variables 
indicated in Table \ref{tab:tab1aidall}.

The median (min-max) follow-up is 
`r pdata %>% summarise(med = dF(median(sos_outtime_death / 365.25 * 12), dig = 1),
                                             min = dF(min(sos_outtime_death / 365.25 * 12), dig = 1),
                                             max = dF(max(sos_outtime_death / 365.25 * 12), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r pdata %>% summarise(sumpy = dF(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.


```{r, child = "../src/km_aid.Rmd"}

```

```{r, child = "../src/outtab_aid.Rmd"}

```

\clearpage
\newpage

# FCM 

In this analysis patients are included if the are ID 
(ferritin <100 µg/L, or ferritin between 100–299 µg/L and transferrin saturation <20%) or
are taking FCM.

## Baseline characteristics according to the use of FCM 

```{r, child = "../src/tab1_fcm.Rmd"}

```

```{r, child = "../src/densitydose_fcm.Rmd"}

```

```{r, child = "../src/geo_fcm.Rmd"}

```

\clearpage
\newpage

## Propensity scores

A propensity score for treatment with FCM was estimated for each patient with 
logistic regression for each of the 10 imputed datasets using the variables 
indicated in Table \ref{tab:tab1fcm}. 1:1 matching without replacement [@match] 
was thereafter performed on the average of the resulting 10 ps [@psmatch]. Matching was 
allowed if the propensity score differed by 0.01 or less. The ability of the ps 
matching to balance the baseline characteristics was assessed by 
standardized mean differences.  

1:1 has deemed the best option when taking into account the number of patients retained.
Other matching ratios: `r matchingn`.

## Outcomes according to the use of FCM

Data were censored at 2018-12-31 or death/emigration. 

The outcomes are presented with the 1 - Kaplan-Meier curves
and tested with the log-rank test. Cox proportional hazards regressions were 
used to model the time to first event analysis and negative binomial regression to model 
the repeated events. The models were partly crude, partly adjusted for 
the individual variables indicated in Table \ref{tab:tab1fcm} and partly in matched cohort, 
modelling the dependence between pairs with a strata term. Note to self: AND WHAT TO DO ABOUT THE NB MODEL????

The median (min-max) follow-up is 
`r pdata_fcm %>% summarise(med = dF(median(sos_outtime_death / 365.25 * 12), dig = 1),
                                             min = dF(min(sos_outtime_death / 365.25 * 12), dig = 1),
                                             max = dF(max(sos_outtime_death / 365.25 * 12), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r pdata_fcm %>% summarise(sumpy = dF(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.


```{r, child = "../src/km_fcm.Rmd"}

```

```{r, child = "../src/outtab_fcm.Rmd"}

```

```{r, child = "../src/outtab_fcm_ef.Rmd"}

```

\clearpage
\newpage

```{r, child = "../src/pred_fcm.Rmd"}

```

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/iron. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
